An Algorithm for Searching Evolving Graphs in Linear Algebraic Language

### 2016-01-23 paper.tex
add adjacency matrices subsection

>>>>

\subsection{Temporal Paths and Adjacency Matrices}
\label{sec:temp-paths-adjac}

%walks on graphs
Let $A$ be the corresponding adjacency matrix of a graph $G$. 
It is a well known result that the $(i,j)$ entry of $A^k$ counts the number of 
walks of length exactly $k$ between node $i$ and node $j$. 
It is natural to consider a generalization of this relationship to 
the case of evolving graphs. 
One may conclude that  the $(i,j)$ entry of the sum of all the product 
\begin{equation}
\label{eq:sum}
A^{[1]}A^{[t_i]}A^{[t_j]} \cdots A^{t_k]}A^{[n]},  \quad t_i \leq t_j \leq \cdots \leq t_k,
\end{equation}
counts the number of temporal paths from $(i,1)$ to $(j,n)$.
However, if we look at it carefully. We observe some temporal paths are 
missed. 
In Figure \ref{fig:eg_shortest_path}, the $(1,3)$ entry of the sum of the product
\[
A^{[1]}A^{[2]}A^{[3]} + A^{[1]}A^{[3]}
\]
is equal to $1$ even though
there are two temporal paths from $(1, t_1)$ to $(3, t_3)$ as shown in Figure \ref{fig:active}.
% Indeed, Grindrod et al. \cite{gphe11} observe
% that the $(i,j)$ entry of $A^{[1]}A^{[2]}\cdots A^{[n]}$ counts the number of 
% temporal paths of length $n$ from $(i, 1)$ to $(j, n)$, which
% go through exactly one edge at each time stamp between 
% $t_1$ and $t_n$. 
% Intuitively, it seems that the $(i,j)$ entry of 
% the summation of all the products of the form
% counts the number of all temporal paths between $i$ and $j$ at all 
% time stamps.  
% However, this is not true since not all temporal paths go through
% exactly one edge at each time stamp. 
% In Figure , the summation of all the products of adjacency matrices
% $A^{[1]}$, $A^{[2]}$, and $A^{[3]}$ 
% \[
% A^{[1]}A^{[2]}A^{[3]}  = 0
% \]
% The $(1,3)$ entry is equal to $1$.  But there are two temporal
% paths from $(1, t_1)$ to $(3, t_3)$ as shown in Figure \ref{fig:active}. 
% missing components call it E
% For adjacency matrices $A^{[s]}$ and $A^{[t]}$, where $s < t$, a nonzero entry
% $(i,j)$ of $A^{[s]} A^{[t]}$ indicates a temporal path from
 $(i,s)$ to $(j,t)$ via
% some node $k$ (which is unknown).
% In general, a nonzero entry $(i,j)$ of the matrix $A^{[1]}A^{[2]}\cdots A^{[n]}$
% indicates a temporal path $p$ that starts at $(i,1)$ and ends at  $(j,n)$.
% Notice the fact that $p$ must go through at least one edge at each time stamp between $1$ and $n$.
Notice $A^{[1]}A^{[2]} = 0$ since 
there is no temporal path from time stamp $t_1$ to $t_2$ that
go through at least one edge at $t_1$. However,
\[
\langle(1, t_1), (1, t_2), (3,t_2)\rangle
\]
is a temporal path as shown in Figure \ref{fig:active}, suggesting we miss to count a temporal path from $(1, t_1)$ to $(3, t_3)$.
Notice also that if we introduce ones on the diagonal of the matrices $\{A^{[i]}\}$,
each node of the resulting evolving graph will become
active at all timestamps.
Hence matrix-matrix multiplication between consecutive
adjacency matrices can not discover all possible temporal paths and should be avoided
when temporal paths are the essential parts of the algorithm.
